Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-11-11 paper/type-inference.tex
Type Inference.
Edits.

Region type templates are FGJ types extended with fresh region
variables ($\rho$) and predicate variables ($\varphi$) to denote
unknown region annotations and region constraints, respectively. For
instance, if a variable $x$ has type $\ObjZ$ in FGJ, its region type
template is of form $\ObjZ\inang{\rho_0}$, where $\rho_0$ is a fresh
region variable. Likewise, given the region-annotated definition of
\C{Pair} class from \S\ref{sec:fb-syntax} a region type template for
the method \C{mkSnd} shown below is \footnote{In our
exposition, we assume that classes $\C{A}$ and $\C{B}$ are trivial
subclasses of $\ObjZ$ with no fields/methods. Like $\ObjZ$, they
accept one region parameter - the allocation region of their
objects.}\footnote{We abuse arrow notation to also represent types of
methods, but unlike function types, there is no allocation region
annotation atop the arrow in a method type.}
$\inang{\rhoalloc_0,\rho_1,\rho_2,\rho_3 \,|\, \varphi_0}
\C{Pair}\inang{\C{A},\C{B}} \inang{\rho_1,\rho_2,\rho_3} \rightarrow
\unitZ$, where $\rhoalloc_0$ and $\rho_{1-3}$ are fresh region
variables, and $\varphi_0$ is a fresh predicate variable denoting
unknown constraints over $\rhoalloc_0$ and $\rho_{1-3}$.  
\begin{codejava}
  unit mkSnd(Pair<A,B> p) {
    return (p.snd = new B());
  }
\end{codejava}
Region type template of a type variable $\C{a}$ is $\C{a}@\rho$, where
$\rho$ is fresh. For a class, a template is computed in two steps. In
the first step, we templatize the types of all its fields, constructor
arguments, and arguments and return values of all its methods, along
with the type of its superclass. In the second step, we generalize all
the free region variables occuring in the templates computed in the
first step as region parameters of the class. Finally, we add a fresh
allocation region parameter ($\rhoalloc$) to the list of parameters,
and introduce a new predicate variable ($\varphi$) to denote unknown
constraints on region parameters. For example, consider the standard
FGJ definition of $\C{Pair}\inang{\C{a},\C{b}}$ class, where $\C{a}$
and $\C{b}$ are the types of $\C{fst}$ and $\C{snd}$ fields,
respectively. It can be templatized as following:
\begin{center}
\begin{codejava}
class Pair$\inang{a\extends\ObjZ, b\extends\ObjZ}$
          $\inang{\rhoalloc_0,\rho_{0-4}\,|\,\varphi_0} \extends \ObjZ\inang{\rho_4}$ {
  $a@\rho_0$ fst;
  $b@\rho_1$ snd;
  Pair($a@\rho_2$ fst, $b@\rho_3$ snd) { $...$ }
  $...$
}
\end{codejava}
\end{center}
% \footnote{In reality, the generalized region variables only come
% from the type templates of class's fields and superclass. Method
% types do not contain any free region variables.}
The template elides $\C{Pair}$'s methods, whose region type templates
contain no free variables. Among the region parameters of the class
template, $\rho_{0-4}$ are obtained by generalizing free region
variables in the types of its class's fields, constructor arguments,
and its superclass type.  The remaining parameter ($\rhoalloc_0$) is a
fresh region variable denoting the allocation region argument. Fresh
predicate variable $\varphi_0$ denotes unknown constraints over
$\rhoalloc_0$ and $\rho_{0-4}$ that need to hold for template to be a
well-formed region-annotated class definition in $\FB$.

@@

We first describe how we solve a non-recursive CSP $\csp$ where
c$\lbrack \varphi,\rhobar \rbrack$ is of the form $\isvalid{\Phicx
\conj \varphi}{\Phics}$.

>>>>

Region type templates are FGJ types extended with fresh region
variables ($\rho$) and predicate variables ($\varphi$) to denote
unknown region annotations and region constraints, respectively. For
instance, if a variable $x$ has type $\ObjZ$ in FGJ, its region type
template is of form $\ObjZ\inang{\rho_0}$, where $\rho_0$ is a fresh
region variable. Likewise, given the region-annotated definition of
\C{Pair} class from \S\ref{sec:fb-syntax} a region type template for a
method with FGJ type $\C{Pair}\inang{\C{A},\C{B}} \rightarrow \C{A}$
is \footnote{In our exposition, we assume that classes $\C{A}$ and
$\C{B}$ are trivial subclasses of $\ObjZ$ with no fields/methods. Like
$\ObjZ$, they accept one region parameter - the allocation region of
their objects.}\footnote{We abuse arrow notation to also represent
types of methods, but unlike function types, there is no allocation
region annotation atop the arrow in a method type.}
$\inang{\rhoalloc_0,\rho_1,\rho_2,\rho_3 \,|\, \varphi_0}
\C{Pair}\inang{\C{A},\C{B}} \inang{\rho_1,\rho_2,\rho_3} \rightarrow
\unitZ$, where $\rhoalloc_0$ and $\rho_{1-3}$ are fresh region
variables, and $\varphi_0$ is a fresh predicate variable denoting
unknown constraints over $\rhoalloc_0$ and $\rho_{1-3}$.  Region type
template of a type variable $\C{a}$ is $\C{a}@\rho$, where $\rho$ is
fresh. For a class, a template is computed in two steps. In the first
step, we templatize the types of all its fields, constructor
arguments, and arguments and return values of all its methods, along
with the type of its superclass. In the second step, we generalize all
the free region variables occuring in the templates computed in the
first step as region parameters of the class. Finally, we add a fresh
allocation region parameter ($\rhoalloc$) to the list of parameters,
and introduce a new predicate variable ($\varphi$) to denote unknown
constraints on region parameters. For example, consider the standard
FGJ definition of $\C{Pair}\inang{\C{a},\C{b}}$ class, where $\C{a}$
and $\C{b}$ are the types of $\C{fst}$ and $\C{snd}$ fields,
respectively. It can be templatized as following:
\begin{center}
\begin{codejava}
class Pair$\inang{a\extends\ObjZ, b\extends\ObjZ}$
          $\inang{\rhoalloc_0,\rho_{0-4}\,|\,\varphi_0} \extends \ObjZ\inang{\rho_4}$ {
  $a@\rho_0$ fst;
  $b@\rho_1$ snd;
  Pair($a@\rho_2$ fst, $b@\rho_3$ snd) { $...$ }
  $...$
}
\end{codejava}
\end{center}
% \footnote{In reality, the generalized region variables only come
% from the type templates of class's fields and superclass. Method
% types do not contain any free region variables.}
The template elides $\C{Pair}$'s methods, whose region type templates
contain no free variables. Among the region parameters of the class
template, $\rho_{0-4}$ are obtained by generalizing free region
variables in the types of its class's fields, constructor arguments,
and its superclass type.  The remaining parameter ($\rhoalloc_0$) is a
fresh region variable denoting the allocation region argument. Fresh
predicate variable $\varphi_0$ denotes unknown constraints over
$\rhoalloc_0$ and $\rho_{0-4}$ that need to hold for template to be a
well-formed region-annotated class definition in $\FB$.

@@

We first describe how we solve a non-recursive CSP $\csp$ where
$c\lbrack \varphi,\rhobar \rbrack$ is of the form $\isvalid{\Phicx
\conj \varphi}{\Phics}$.
